% backmatter/appendix_questionnaire.tex
% This appendix contains the sample questionnaire used for the household survey.

\chapter{Sample Household Survey Questionnaire}
\label{app:questionnaire}

\textbf{Questionnaire ID:} \hrulefill \hspace{1cm}
\textbf{Date of Interview:} \hrulefill \\[1em]

\section*{Section 1: Socio-Demographic Information}

\begin{enumerate}[label=\arabic*.]
    \item Gender of Respondent:
    \begin{itemize}
        \item[$\square$] Male
        \item[$\square$] Female
    \end{itemize}

    \item Age of Respondent (Years): \hrulefill

    \item Household Size (Number of persons): \hrulefill

    \item Education Level of Respondent:
    \begin{itemize}
        \item[$\square$] No formal education
        \item[$\square$] Primary education (Grades 1--8)
        \item[$\square$] Secondary education (Grades 9--12)
        \item[$\square$] Tertiary education (College/University)
    \end{itemize}

    \item Main Occupation of Household Head: \hrulefill

    \item Approximate Monthly Household Income (Placeholder Currency): \hrulefill

    \item How long have you lived in this area (years)? \hrulefill

    \item Approximate distance of your house from the disposal site:
    \begin{itemize}
        \item[$\square$] < 500 meters
        \item[$\square$] 500 meters -- 1 km
        \item[$\square$] > 1 km \quad (If > 1 km, thank and end interview)
    \end{itemize}
\end{enumerate}

\section*{Section 2: Solid Waste Management Practices}

\begin{enumerate}[label=\arabic*., start=9]
    \item How do you dispose of your household waste? (Select all that apply)
    \begin{itemize}
        \item[$\square$] Municipal collection (door-to-door)
        \item[$\square$] Municipal collection (communal bins)
        \item[$\square$] Open dumping (near house/in open field)
        \item[$\square$] Burning
        \item[$\square$] Burying
        \item[$\square$] Given to waste pickers
        \item[$\square$] Other (specify): \hrulefill
    \end{itemize}

    \item How often is waste collected from your area by the municipality (if applicable)?
    \begin{itemize}
        \item[$\square$] Daily
        \item[$\square$] 2--3 times a week
        \item[$\square$] Once a week
        \item[$\square$] Less than once a week
        \item[$\square$] Never
    \end{itemize}

    \item Do you segregate your waste at home (e.g., organic, plastic, paper)?
    \begin{itemize}
        \item[$\square$] Yes \quad \item[$\square$] No
    \end{itemize}
    If Yes, how? \hrulefill

    \item What challenges do you face in disposing of your waste properly? \\
    \vspace{1em}\hrulefill \\[1em]
    \hrulefill
\end{enumerate}

\section*{Section 3: Health Status and Perceptions}

\begin{enumerate}[label=\arabic*., start=13]
    \item In the past six months, has any member of your household experienced the following health problems? (Tick all that apply)
    \begin{itemize}
        \item[$\square$] Acute Respiratory Infections (persistent cough, difficulty breathing)
        \item[$\square$] Diarrheal Diseases (cholera, typhoid, loose stools)
        \item[$\square$] Skin Rashes/Irritations
        \item[$\square$] Eye Irritations
        \item[$\square$] Malaria or other vector-borne diseases (dengue, typhus)
        \item[$\square$] Headaches/Nausea
        \item[$\square$] Other (specify): \hrulefill
        \item[$\square$] None of the above
    \end{itemize}

    \item Do you believe there is a link between your family’s health problems and the solid waste disposal site?
    \begin{itemize}
        \item[$\square$] Yes
        \item[$\square$] No
        \item[$\square$] Don’t know
    \end{itemize}

    \item If Yes, please explain: \\
    \vspace{1em}\hrulefill \\[1em]
    \hrulefill

    \item Are you aware of the health risks associated with improper solid waste disposal?
    \begin{itemize}
        \item[$\square$] Yes \quad \item[$\square$] No
    \end{itemize}

    \item What are your attitudes towards improving SWM?
    \begin{itemize}
        \item[$\square$] Very positive
        \item[$\square$] Positive
        \item[$\square$] Neutral
        \item[$\square$] Negative
        \item[$\square$] Very negative
    \end{itemize}

    \item Would you be willing to pay a small fee for improved waste collection and disposal services?
    \begin{itemize}
        \item[$\square$] Yes
        \item[$\square$] No
        \item[$\square$] Maybe
    \end{itemize}

    \item Do you feel that public health education on waste management is sufficient in your community?
    \begin{itemize}
        \item[$\square$] Yes
        \item[$\square$] No
    \end{itemize}

    \item What suggestions do you have for improving SWM and protecting public health in your area? \\
    \vspace{1em}\hrulefill \\[1em]
    \hrulefill
\end{enumerate}

\bigskip
\textbf{Thank you for your participation!}

